The Unicode Cookbook \linebreak for Linguists

### 2018-06-22 book/chapters/implementation.tex
gets code display correctling in implementation chapter

\begin{lstlisting}[language=bash]
  $ pip install segments
\end{lstlisting}

@@

\begin{lstlisting}[language=bash]
  $ git clone https://github.com/cldf/segments
  $ cd segments
  $ python setup.py develop
\end{lstlisting}

@@

\begin{lstlisting}
>>> from segments.tokenizer import Tokenizer
\end{lstlisting}

@@

% \lstset{extendedchars=false, escapeinside=**}
\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> t = Tokenizer()
>>> result = t.characters('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)')
>>> print(result)
>>> '(*@c ̂ h a ́ ɾ a ̃ ̌ c t ʼ ɛ ↗ ʐ ː | # k ͡ p@*)'
\end{lstlisting}

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> result = t.grapheme_clusters('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)')
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c t ʼ ɛ ↗ ʐ ː | # k͡ p@*)'
\end{lstlisting}

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> result = t('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)')
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c t ʼ ɛ ↗ ʐ ː | # k͡ p@*)'
\end{lstlisting}

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> result = t('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)', ipa=True)
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c tʼ ɛ ↗ ʐː | # k͡p@*)'
\end{lstlisting}

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
  $ more data/orthography-profile.tsv
  Grapheme  IPA   XSAMPA  COMMENT
  a         a     a
  aa        (*@aː@*)    (*@a:@*)
  b         b     b
  c         c     c
  ch        (*@tʃ@*)    tS
  -         NULL  NULL    "comment with   tab"
  on        (*@õ@*)     o~
  n         n     n
  ih        (*@í@*)     i_H
  inh       (*@ĩ́@*)     i~_H
\end{lstlisting}

@@

\begin{lstlisting}
>>> from segments.tokenizer import Profile
>>> t = Tokenizer('data/orthography-profile.tsv')
\end{lstlisting}

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> t('(*@aabchonn-ih@*)')
>>> '(*@aa b ch on n - ih@*)'
\end{lstlisting}

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> t.transform('(*@aabchonn-ih@*)', 'IPA')
>>> '(*@aː b tʃ õ n í@*)'
\end{lstlisting}

@@

\begin{lstlisting}
>>> t.transform('aabchonn-ih', 'XSAMPA')
>>> 'a: b tS o~ n i_H'
\end{lstlisting}

@@

\begin{lstlisting}
>>> t = Tokenizer('data/orthography-profile.tsv', 
	errors_replace=lambda c: '?')
>>> t.find_missing_characters("aa b ch on n - ih x y z")
>>> 'aa b ch on n - ih ? ? ?'
\end{lstlisting}

\begin{lstlisting}
>>> t = Tokenizer('data/orthography-profile.tsv', 
	errors_replace=lambda c: '<{0}>'.format(c))
>>> t.find_missing_characters("aa b ch on n - ih x y z")
>>> 'aa b ch on n - ih <x> <y> <z>'
\end{lstlisting}

@@

\begin{lstlisting}
>>> profile = Profile.from_text("aa b ch on n - ih x y z")
>>> print(profile)
\end{lstlisting}

\begin{lstlisting}[language=bash,texcl=true]
  Grapheme  frequency  mapping
            9
  a         2          a
  h         2          h
  n         2          n
  b         1          b
  c         1          c
  o         1          o
  -         1          -
  i         1          i
  x         1          x
  y         1          y
  z         1          z
\end{lstlisting}

@@

\begin{lstlisting}[language=bash]
  $ segments -h

usage: segments [-h] [--verbosity VERBOSITY] 
                     [--encoding ENCODING]
                     [--profile PROFILE]
                     [--mapping MAPPING]
                     command ...

@@

optional arguments:
  -h, --help            show this help message and exit
  --verbosity VERBOSITY
                        increase output verbosity
  --encoding ENCODING   input encoding
  --profile PROFILE     path to an orthography profile
  --mapping MAPPING     column name in ortho profile to map 
                        graphemes

@@

\begin{lstlisting}[language=bash]
  $ more sources/german.txt

@@

\begin{lstlisting}[language=bash, extendedchars=false, escapeinside={(*@}{@*)}]
  $ cat sources/german.txt | segments profile

@@

\begin{lstlisting}[language=bash, extendedchars=false, escapeinside={(*@}{@*)}]
  $ more more data/german-orthography-profile.tsv

@@

\begin{lstlisting}[language=bash, extendedchars=false, escapeinside={(*@}{@*)}]
  $ cat sources/german.txt | segments 
    --profile=data/german-orthography-profile.tsv tokenize

@@

\begin{lstlisting}[language=bash, extendedchars=false, escapeinside={(*@}{@*)}]
  $ cat sources/german.txt | segments --mapping=IPA 
	--profile=data/german-orthography-profile.tsv tokenize

@@

\begin{lstlisting}[language=bash]
  $ cat sources/german.txt | segments --mapping=XSAMPA 
	--profile=data/german-orthography-profile.tsv tokenize

@@

% Old example that reviewer didn't like!
\begin{comment}
\begin{lstlisting}[language=bash]
  $ more text.txt
  aäaaöaaüaa
\end{lstlisting}

@@

  $ cat text.txt | segments --mapping=mapping 
    --profile=profile.prf tokenize
  a ä x ö x ü x
\end{lstlisting}
\end{comment}
% \end old example that reviewer didn't like!

>>>>

\begin{lstlisting}[language=bash, basicstyle=\myfont]
  $ pip install segments
\end{lstlisting}

@@

\begin{lstlisting}[language=bash, basicstyle=\myfont]
  $ git clone https://github.com/cldf/segments
  $ cd segments
  $ python setup.py develop
\end{lstlisting}

@@

\begin{lstlisting}[basicstyle=\myfont]
>>> from segments.tokenizer import Tokenizer
\end{lstlisting}

@@

% \lstset{extendedchars=false, escapeinside=**}
\begin{lstlisting}[basicstyle=\myfont, extendedchars=false, escapeinside={(*@}{@*)}]
>>> t = Tokenizer()
>>> result = t.characters('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)')
>>> print(result)
>>> '(*@c ̂ h a ́ ɾ a ̃ ̌ c t ʼ ɛ ↗ ʐ ː | # k ͡ p@*)'
\end{lstlisting}

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont]
>>> result = t.grapheme_clusters('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)')
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c t ʼ ɛ ↗ ʐ ː | # k͡ p@*)'
\end{lstlisting}

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont]
>>> result = t('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)')
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c t ʼ ɛ ↗ ʐ ː | # k͡ p@*)'
\end{lstlisting}

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont]
>>> result = t('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)', ipa=True)
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c tʼ ɛ ↗ ʐː | # k͡p@*)'
\end{lstlisting}

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont, showstringspaces=false]
  $ more data/orthography-profile.tsv
  Grapheme  IPA   XSAMPA  COMMENT
  a         a     a
  aa        (*@aː@*)    (*@a:@*)
  b         b     b
  c         c     c
  ch        (*@tʃ@*)    tS
  (*@-@*)         NULL  NULL    "comment with   tab"
  on        (*@õ@*)     o~
  n         n     n
  ih        (*@í@*)     i_H
  inh       (*@ĩ́@*)     i~_H
\end{lstlisting}

@@

\begin{lstlisting}[basicstyle=\myfont]
>>> from segments.tokenizer import Profile
>>> t = Tokenizer('data/orthography-profile.tsv')
\end{lstlisting}

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont]
>>> t('(*@aabchonn-ih@*)')
>>> '(*@aa b ch on n - ih@*)'
\end{lstlisting}

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont]
>>> t.transform('(*@aabchonn-ih@*)', 'IPA')
>>> '(*@aː b tʃ õ n í@*)'
\end{lstlisting}

@@

\begin{lstlisting}[basicstyle=\myfont, showstringspaces=false]
>>> t.transform('aabchonn-ih', 'XSAMPA')
>>> 'a: b tS o~ n i_H'
\end{lstlisting}

@@

\begin{lstlisting}[basicstyle=\myfont, extendedchars=false, escapeinside={(*@}{@*)}, showstringspaces=false]
>>> t = Tokenizer('data/orthography(*@-@*)profile.tsv', 
	errors_replace=lambda c: '?')
>>> t.find_missing_characters("aa b ch on n (*@-@*) ih x y z")
>>> 'aa b ch on n (*@-@*) ih ? ? ?'
\end{lstlisting}

\begin{lstlisting}[basicstyle=\myfont, extendedchars=false, escapeinside={(*@}{@*)}, showstringspaces=false]
>>> t = Tokenizer('data/orthography(*@-@*)profile.tsv', 
	errors_replace=lambda c: '<{0}>'.format(c))
>>> t.find_missing_characters("aa b ch on n (*@-@*) ih x y z")
>>> 'aa b ch on n (*@-@*) ih <x> <y> <z>'
\end{lstlisting}

@@

\begin{lstlisting}[basicstyle=\myfont, extendedchars=false, escapeinside={(*@}{@*)}, showstringspaces=false]
>>> profile = Profile.from_text("aa b ch on n (*@-@*) ih x y z")
>>> print(profile)
\end{lstlisting}

\begin{lstlisting}[language=bash, texcl=true, basicstyle=\myfont, extendedchars=false, escapeinside={(*@}{@*)}]
  Grapheme  frequency  mapping
            9
  a         2          a
  h         2          h
  n         2          n
  b         1          b
  c         1          c
  o         1          o
  (*@-@*)         1          (*@-@*)
  i         1          i
  x         1          x
  y         1          y
  z         1          z
\end{lstlisting}

@@

\begin{lstlisting}[language=bash, basicstyle=\myfont, extendedchars=false, escapeinside={(*@}{@*)}]
  $ segments (*@-@*)h

usage: segments [(*@-@*)h] [(*@--@*)verbosity VERBOSITY] 
                     [(*@--@*)encoding ENCODING]
                     [(*@--@*)profile PROFILE]
                     [(*@--@*)mapping MAPPING]
                     command ...

@@

optional arguments:
  (*@-@*)h, (*@--@*)help            show this help message and exit
  (*@--@*)verbosity VERBOSITY
                        increase output verbosity
  (*@--@*)encoding ENCODING   input encoding
  (*@--@*)profile PROFILE     path to an orthography profile
  (*@--@*)mapping MAPPING     column name in ortho profile to map 
                        graphemes

@@

\begin{lstlisting}[language=bash, basicstyle=\myfont]
  $ more sources/german.txt

@@

\begin{lstlisting}[language=bash, extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont]
  $ cat sources/german.txt | segments profile

@@

\begin{lstlisting}[language=bash, extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont]
  $ more data/german-orthography-profile.tsv

@@

\begin{lstlisting}[language=bash, extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont]
  $ cat sources/german.txt | segments 
    (*@--@*)profile=data/german(*@-@*)orthography(*@-@*)profile.tsv tokenize

@@

\begin{lstlisting}[language=bash, extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont]
  $ cat sources/german.txt | segments (*@--@*)mapping=IPA 
	(*@--@*)profile=data/german(*@-@*)orthography(*@-@*)profile.tsv tokenize

@@

\begin{lstlisting}[language=bash, basicstyle=\myfont, extendedchars=false, escapeinside={(*@}{@*)}]
  $ cat sources/german.txt | segments (*@--@*)mapping=XSAMPA 
	(*@--@*)profile=data/german(*@-orthography(*@-@*)profile.tsv tokenize

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Old example that reviewer didn't like!
\begin{comment}
\begin{lstlisting}[language=bash]
  $ more text.txt
  aäaaöaaüaa
\end{lstlisting}

@@

  $ cat text.txt | segments --mapping=mapping 
    --profile=profile.prf tokenize
  a ä x ö x ü x
\end{lstlisting}
\end{comment}
% \end old example that reviewer didn't like!
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
